Open source landscape ecology tools \thanks{Grants or other notes about the article that should go on the front page should be placed here. General acknowledgments should be placed at the end of the article.}

### 2020-10-15 paper/paper.Rmd
updates ref

The use of landscape metrics was heavily facilitated by the introduction of the *FRAGSTATS* software \cite{McGarigal2012}, firstly published in 1995 \cite{Kupfer2012,Gustafson2019}.
However, *FRAGSTATS* is not not open-source software (however it is free to download at the developers' homepage), which restricts transparency, reproducibility, and extensibility.
To provide a truly open-source solution to calculate landscape metrics, recently the *landscapemetrics* package \cite{Hesselbarth2019a} was developed.
The package allows to calculate the most widely used landscape metrics in a transparent and reproducible workflow within the *R* environment. 
Additionally, the *belg* packages allows to calculate the Boltzmann entropy of a landscape gradient \cite{Nowosad2019a}.
<!-- JN: two notes: -->
<!-- 1. I am working on a paper supporting the belg package - I plan to submit it in less than a month, so hopefully we could update the reference here -->
<!-- 2. I also submitted a paper yesterday (2020-07-17) explaining a new package motif, which is aimed on calculations based on spatial signatures. I plan to add a paragraph about this package someplace in this paper. -->
<!-- MH: Sounds great -->

@@

  \begin{table}
    \begin{tabularx}{0.9\linewidth}{XXXX}
      \caption{Overview of commonly used \textit{R} packages for landscape ecology. Packages are sorted by their major application topic. Only packages especially for spatial data and landscape ecology are included.} \\
      \hline
      Topic & Package & Description & Source \\
      \hline
      \multirow{5}{*}{Spatial data} & \textit{raster} & Methods for raster data & \cite{Hijmans2019} \\
      & \textit{stars} & Methods for spatiotemporal arrays (raster and vector data cubes) & \cite{Pebesma2019} \\
      & \textit{sp} & Methods for mainly vector data & \cite{Pebesma2005,Bivand2013} \\
      & \textit{sf} & Methods for vector data as simple features & \cite{Pebesma2018} \\
      & \textit{landscapetools} & Utility functions for landscape analysis & \cite{Sciaini2018} \\
      \hline
      \multirow{3}{*}{Creating maps} & \textit{tmap} & xxx & xxx \\
      & \textit{mapview} & xxx & xxx \\
      & \textit{leaflet} & xxx & xxx \\
      \hline
      \multirow{2}{*}{Landscape metrics} & \textit{landscapemetrics} & Calculates landscape metrics for categorical land cover maps & \cite{Hesselbarth2019a} \\
      & \textit{belg} & Calculates the Boltzmann entropy of a landscape gradient & \cite{Nowosad2019a} \\
      \hline
      \multirow{5}{*}{}Species distribution modeling (SDM) & \textit{dismo} & xxx & xxx \\
      & \textit{sdm} & xxx & xxx \\
      & \textit{ecospat} & xxx & xxx \\
      & \textit{biomod2} & xxx & xxx \\
      & \textit{PresenceAbsence} & xxx & xxx \\
      \hline
      \multirow{3}{*}{Connectivity} & \textit{grainscape} & xxx & \cite{Chubaty2020} \\
      & \textit{gdistance} & xxx & \cite{vanEtten2017} \\
      & \textit{samc} & xxx & \cite{Marx2020} \\
      \hline
      \multirow{3}{*}{Landscape genetics} & \textit{graphs4lg} & xxx & \cite{Savary2020} \\
      & \textit{PopGenReport} & xxx & \cite{Adamack2014,Gruber2015} \\
      & \textit{HierDpart} & xxx & \cite{Qin2019} \\
      \hline
      Neutral Landscape Models (NLM) & \textit{NLMR} & xxx & \cite{Sciaini2018} \\
      \hline
    \end{tabularx}
  \end{table}

>>>>

The use of landscape metrics was heavily facilitated by the introduction of the *FRAGSTATS* software \cite{McGarigal2012}, firstly published in 1995 \cite{Kupfer2012,Gustafson2019}.
However, *FRAGSTATS* is not open-source software (however it is free to download at the developers' homepage), which restricts transparency, reproducibility, and extensibility.
To provide a truly open-source solution to calculate landscape metrics, recently the *landscapemetrics* package \cite{Hesselbarth2019a} was developed.
The package allows to calculate the most widely used landscape metrics in a transparent and reproducible workflow within the *R* environment. 
Additionally, the *belg* packages allows to calculate the Boltzmann entropy of a landscape gradient \cite{R-belgpaper}.
<!-- JN: two notes: -->
<!-- 1. I am working on a paper supporting the belg package - I plan to submit it in less than a month, so hopefully we could update the reference here -->
<!-- 2. I also submitted a paper yesterday (2020-07-17) explaining a new package motif, which is aimed on calculations based on spatial signatures. I plan to add a paragraph about this package someplace in this paper. -->
<!-- MH: Sounds great -->

@@

  \begin{table}
    \begin{tabularx}{0.9\linewidth}{XXXX}
      \caption{Overview of commonly used \textit{R} packages for landscape ecology. Packages are sorted by their major application topic. Only packages especially for spatial data and landscape ecology are included.} \\
      \hline
      Topic & Package & Description & Source \\
      \hline
      \multirow{5}{*}{Spatial data} & \textit{raster} & Methods for raster data & \cite{Hijmans2019} \\
      & \textit{stars} & Methods for spatiotemporal arrays (raster and vector data cubes) & \cite{Pebesma2019} \\
      & \textit{sp} & Methods for mainly vector data & \cite{Pebesma2005,Bivand2013} \\
      & \textit{sf} & Methods for vector data as simple features & \cite{Pebesma2018} \\
      & \textit{landscapetools} & Utility functions for landscape analysis & \cite{Sciaini2018} \\
      \hline
      \multirow{3}{*}{Creating maps} & \textit{tmap} & xxx & xxx \\
      & \textit{mapview} & xxx & xxx \\
      & \textit{leaflet} & xxx & xxx \\
      \hline
      \multirow{2}{*}{Landscape metrics} & \textit{landscapemetrics} & Calculates landscape metrics for categorical land cover maps & \cite{Hesselbarth2019a} \\
      & \textit{belg} & Calculates the Boltzmann entropy of a landscape gradient & \cite{R-belgpaper} \\
      \hline
      \multirow{5}{*}{}Species distribution modeling (SDM) & \textit{dismo} & xxx & xxx \\
      & \textit{sdm} & xxx & xxx \\
      & \textit{ecospat} & xxx & xxx \\
      & \textit{biomod2} & xxx & xxx \\
      & \textit{PresenceAbsence} & xxx & xxx \\
      \hline
      \multirow{3}{*}{Connectivity} & \textit{grainscape} & xxx & \cite{Chubaty2020} \\
      & \textit{gdistance} & xxx & \cite{vanEtten2017} \\
      & \textit{samc} & xxx & \cite{Marx2020} \\
      \hline
      \multirow{3}{*}{Landscape genetics} & \textit{graphs4lg} & xxx & \cite{Savary2020} \\
      & \textit{PopGenReport} & xxx & \cite{Adamack2014,Gruber2015} \\
      & \textit{HierDpart} & xxx & \cite{Qin2019} \\
      \hline
      Neutral Landscape Models (NLM) & \textit{NLMR} & xxx & \cite{Sciaini2018} \\
      \hline
    \end{tabularx}
  \end{table}

### 2020-10-15 paper/paper.tex
updates ref

The use of landscape metrics was heavily facilitated by the introduction of the \emph{FRAGSTATS} software \cite{McGarigal2012}, firstly published in 1995 \cite{Kupfer2012,Gustafson2019}.
However, \emph{FRAGSTATS} is not not open-source software (however it is free to download at the developers' homepage), which restricts transparency, reproducibility, and extensibility.
To provide a truly open-source solution to calculate landscape metrics, recently the \emph{landscapemetrics} package \cite{Hesselbarth2019a} was developed.
The package allows to calculate the most widely used landscape metrics in a transparent and reproducible workflow within the \emph{R} environment.
Additionally, the \emph{belg} packages allows to calculate the Boltzmann entropy of a landscape gradient \cite{Nowosad2019a}.

>>>>

The use of landscape metrics was heavily facilitated by the introduction of the \emph{FRAGSTATS} software \cite{McGarigal2012}, firstly published in 1995 \cite{Kupfer2012,Gustafson2019}.
However, \emph{FRAGSTATS} is not open-source software (however it is free to download at the developers' homepage), which restricts transparency, reproducibility, and extensibility.
To provide a truly open-source solution to calculate landscape metrics, recently the \emph{landscapemetrics} package \cite{Hesselbarth2019a} was developed.
The package allows to calculate the most widely used landscape metrics in a transparent and reproducible workflow within the \emph{R} environment.
Additionally, the \emph{belg} packages allows to calculate the Boltzmann entropy of a landscape gradient \cite{R-belgpaper}.
